---
title: "Population Health and Economy impact by sever weather events from 1950 to 2012"
author: "Shinkuang Chang"
date: "July 23, 2015"
output: 
  html_document: 
    keep_md: yes
---

#Synopsis

In this report we try to find out which severe weather event is most deadly to population and which severe weather event is most costly to economy accross U.S. This study will use the storm data compiled by National Oceanic and Atmospheric Administration (NOAA). This data covers the severe wether events from 1950 to November of 2011. 


#Data Processing

##Reading the whole data 

```{r load}
# Disable scientific notation
options(scipen=999)

filename <- "repdata_data_StormData.csv"
# Read NOAA severe weather events data to a data frame
dt0 <- read.table(filename,sep=",",header=T,na.strings = "")
```

```{r dimension}
dim(dt0)
num_records <- NROW(dt0)
```
There are `r num_records` in `r filename` dataset. 

## Checking data set variables especially FATALITIES, and PROPDMG for missing data

Show first few rows of records for the first seven variables.

```{r head}
head(dt0[,1:7])
```


Since We are interesting FATALITIES and PROPDMG variables from the dataset, we are retrieving FATALITIES to a dataset called fatality and PROPDMG to a dataset called propdmg. We will use the summary function in R to show some simple summaries of these two datasets. 

```{r getvarialbe}
fatality <- dt0$FATALITIES
summary(fatality)

propdmg <- dt0$PROPDMG
summary(propdmg)
```

From the simple summaries we can see that there are no missing values for both FATALITIES and PROMDMG variables from the original datasets and this is good for the following data analysis. 


#Result

## Entire U.S. analysis

### Find out which severe weather event is most deadly accorss the U.S. from 1950 to November 2011. 

We will sum up all the death records by severe weather event type and sum up all the property damage by the severe weather event type. Then we will find out which severe weather event from both sum of sub dataset has the maximum value for total death and total property damage. 


```{r analysis}
dt1 <- aggregate(FATALITIES ~ EVTYPE, data = dt0, FUN=sum)

event1 <- dt1[which.max(dt1$FATALITIES),1]
total_death <- dt1[which.max(dt1$FATALITIES),2]

dt2 <- aggregate(PROPDMG ~ EVTYPE, data = dt0, FUN = sum)
event2 <- dt2[which.max(dt2$PROPDMG),1]
total_damage <- dt2[which.max(dt2$PROPDMG),2]
```

After the analysis, we find out - **`r event1`** is the most deadly severe weather event accross U.S. from 1950 to November 2011 which has a total **`r total_death`** people dead and **`r event2`** is the most costly severe weather event accross U.S. from 1950 to November 2011 which cause U.S. **`r total_damage`** millions U.S. dollars. 


Following graph will show the top tep deadly severe weather events accross U.S. from 1950 to November 2011 with most casualties. 


```{r}
par(mai=c(1, 1, 1, 1))
dt1_ordered <- dt1[order(-dt1$FATALITIES),]
dt1_top10 <- dt1_ordered[1:10,]
dt1_top10
dt1_xaxis_label <- dt1_top10$EVTYPE
dt1_xaxis_label
barplot(dt1_top10$FATALITIES,main="Top 10 deadly severe weather events", ylab="Total death",names.arg = dt1_xaxis_label, las = 2)

```

Following graph will show the top 10 costly severe weather events accross U.S. from 1950 to November 2011 with millions dollars lost. 


```{r}
par(mai=c(1, 1, 1, 1))
dt2_ordered <- dt2[order(-dt2$PROPDMG),]
dt2_top10 <- dt2_ordered[1:10,]
dt2_top10
dt2_xaxis_label <- dt2_top10$EVTYPE
dt2_xaxis_label
barplot(dt2_top10$PROPDMG,main="Top 10 costly severe weather events damage in millions", names.arg = dt2_xaxis_label, las = 2)

```



