---
title: "Population Health and Economy impact by sever weather events from 1950 to 2012"
author: "Shinkuang Chang"
date: "July 23, 2015"
output: 
  html_document: 
    keep_md: yes
---

#Synopsis
In this report we try to find out which severe weather event is most harmful to population and which severe weather event is most costly to economy through all the states. This study will be based on the storm data compiled by National Oceanic and Atmospheric Administration (NOAA). This data covers the severe wether events from 1950 to November of 2011. 


The data set is coming from ...

#Loading and Processing the Raw Data
##Reading the whole data 

```{r load}
# Disable scientific notation
options(scipen=999)

filename <- "repdata_data_StormData.csv"
# Read NOAA severe weather events data to a data frame
dt0 <- read.table(filename,sep=",",header=T,na.strings = "")
```

```{r dimension}
dim(dt0)
num_records <- NROW(dt0)
```
There are `r num_records` in `r filename` dataset. 

## Checking data set especially FATALITIES, and PROPDMG
Show first few rows of records for the first six variables.

```{r head}
head(dt0[,1:6])
```


We are retrieving the FATALITIES to a dataset called fatality and PROPDMG to a dataset called propdmg. We will use the summary function in R to show some simple summaries of these two datasets. 

```{r getvarialbe}
# show the first few line of records
evtype <- dt0$EVTYPE
fatality <- dt0$FATALITIES
summary(fatality)
propdmg <- dt0$PROPDMG
summary(propdmg)
```

From above code you can see that there are no missing values for both FATALITIES and PROMDMG variables from the original datasets and this is good for the following data analysis. 


#Result
## Entire U.S. analysis
```{r analysis}

dt1 <- aggregate(FATALITIES ~ EVTYPE, data = dt0, FUN=sum)

event1 <- dt1[which.max(dt1$FATALITIES),1]
total_death <- dt1[which.max(dt1$FATALITIES),2]

dt2 <- aggregate(PROPDMG ~ EVTYPE, data = dt0, FUN = sum)
event2 <- dt2[which.max(dt2$PROMDMG),1]
total_damage <- dt2[which.max(dt2$PROPDMG),2]
```

`r event1` is the most deadly severe weather event accross U.S. from 1950 to November 2011. 
Total death for `r event1` is `r total_death`.

`r event2` is the most costly severe weather event accross U.S. from 1950 to November 2011. 
Total damage cost for `r event2` is `r total_damage` millions U.S. dollars. 


Following graph will show the top 10 deadly severe weather events accross U.S. from 1950 to November 2011.


```{r}
par(mai=c(1, 1, 1, 1))
dt1_ordered <- dt1[order(-dt1$FATALITIES),]
dt1_top10 <- dt1_ordered[1:10,]
dt1_top10
dt1_xaxis_label <- dt1_top10$EVTYPE
dt1_xaxis_label
barplot(dt1_top10$FATALITIES,main="Top 10 deadly severe weather events", ylab="Total death",names.arg = dt1_xaxis_label, las = 2)

```

Following graph will show the top 10 costly severe weather events accross U.S. from 1950 to November 2011.


```{r}
par(mai=c(1, 1, 1, 1))
dt2_ordered <- dt2[order(-dt2$PROPDMG),]
dt2_top10 <- dt2_ordered[1:10,]
dt2_top10
dt2_xaxis_label <- dt2_top10$EVTYPE
dt2_xaxis_label
barplot(dt2_top10$PROPDMG,main="Top 10 costly severe weather events damage in millions", names.arg = dt2_xaxis_label, las = 2)

```



